---
title: Data Exploration
output: html_document
author: Jeffrey Mei
date: 2024-10-26
---

# Data Exploration

```{R}
# Read Data
filepath <- "~/Documents/School/BIOS511/Project/data/owner_change.csv"
owner_change <- read_csv(filepath)

#project_directory <- "~/Documents/School/BIOS511/Project"
project_directory <- "~/Documents/2024.03 - Fall/BIOS511/Final Project/Hospital-Privatization" # replace this with your own project directory

hospital_info_path <- "/data/HOSPITAL_INFO.csv"
hospital_info <- read_csv(paste(project_directory, hospital_info_path, sep = ""))
```

```{R}
# Histogram
ggplot(owner_change, aes(x = `EFFECTIVE DATE`)) +
  geom_histogram(binwidth = 30, color = "black", fill = "lightblue") +
  labs(title = "Histogram of Effective Dates", x = "Effective Date", y = "Count") +
  theme_minimal()
```

```{R}
# Link Data
hospital_changes <-
  left_join(
    owner_change %>% select(
      "CCN - BUYER",
      "CCN - SELLER",
      "ORGANIZATION NAME - BUYER",
      "ORGANIZATION NAME - SELLER",
      "EFFECTIVE DATE"
    ),
    hospital_info %>% select(
      "Facility ID",
      "Facility Name"
    ),
    by = c("CCN - BUYER" = "Facility ID")
  ) %>%
  left_join(
    hospital_info %>% select(
      "Facility ID",
      "Facility Name"
    ),
    by = c("CCN - SELLER" = "Facility ID")
  ) %>%
  # Renaming the Redundant Columns
  rename(
    "Facility Name - Buyer" = "Facility Name.x",
    "Facility Name - Seller" = "Facility Name.y"
  )
```

Let's make sure that we have a proportional sample throughout time. Once we filter out all _incomplete data_ - owner changes without buyer _and_ seller information - is our data missing at random? The following figure appears to support this claim. 

```{R}
hospital_changes %>%
  # Add Temporary Label for Complete Data: Existing Buyer/Seller Info
  mutate(complete = case_when(
    is.na(`Facility Name - Buyer`) |
      is.na(`Facility Name - Seller`) ~ "Missing",
    TRUE ~ "Complete"
  )) %>%
  mutate(complete = factor(complete, levels = c("Missing", "Complete"))) %>%
  # Plot Histogram
  ggplot(
    aes(
      x = `EFFECTIVE DATE`,
      fill = complete,
    )
  ) +
  geom_histogram(
    binwidth = 30,
    color = "black",
    alpha = 0.5,
    position = "stack"
  ) +

  # Set Legend
  scale_fill_manual(
    name = "Data Completeness",
    values = c("Complete" = "red", "Missing" = "gray"),
    labels = c("Complete" = "Complete Data", "Missing" = "Missing Data")
  ) +

  # Additional aesthetics
  labs(
    title = "Hospital Changes Histogram",
    x = "Effective Date",
    y = "Count"
  ) +
  theme_minimal()
```

Pivot the survey data to make it more usable. Each hospital should have one row consisting of all available survey data. Also remove any columns that seem to have no data in them at all. 
```{r}
PES_path <- "/data/PE Hospital Tracker.csv"
survey_path <- "/data/patient_survey/HCAHPS-Hospital.csv"
PES_data <- read_csv(paste(project_directory, PES_path, sep = ""))
survey_data <- read_csv(paste(project_directory, survey_path, sep = ""))
survey_cleaned = survey_data %>%
  select(c("Facility ID", "Facility Name", "HCAHPS Question", "Patient Survey Star Rating", "Start Date", "End Date")) %>%
  pivot_wider(names_from = "HCAHPS Question", values_from =  "Patient Survey Star Rating") %>%
  select(where(~ n_distinct(.) > 1))
```

Join the data and coerce one of the columns to be an integer. Plot the resulting data and we immediately see one issue we're going to run into - there are way more hospitals that are NOT owned by PE. Plus it looks like the averages are not significantly different - at least the data look relatively normal? I will work on modifying this script to look at each of the survey items, though I'm not sure any of them will be any different. 
```{r}
survey_owner_data = left_join(survey_test, PES_data, by = c("Facility ID"="Facility ID (CCN)"))
survey_owner_data$"Nurse communication - star rating" = as.integer(survey_owner_data$`Nurse communication - star rating`)
ggplot(survey_owner_data, aes(x = `Currently PE?`, y =`Nurse communication - star rating`)) +
  geom_jitter(width = 0.2)
t.test(survey_owner_data[survey_owner_data$`Currently PE?` == "Y",3],survey_owner_data[survey_owner_data$`Currently PE?` == "N",3])
hist(as.numeric(unlist(survey_owner_data[!is.na(survey_owner_data[,3]),3])), breaks=0.5+seq(0,5))
```




